CorReg: model-based pretreatment for regression with correlated variables. Application in steel industry

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@479 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
We note the linear regression model:
\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}). As shown in \ref{sectionOLS}, estimation of the parameters of the regression we have to compute the inverse of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\hat{\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model  : the more covariates you have the greater variance you get.
	\item The correlations within the covariates : strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

@@

	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we don't have a system of regressions but one regression on our response variable and a system of sub-regressions (without the response variable). Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss. It can be seen as a pretreatment on the dataset because other regression methods can then be used. So we may obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.

>>>>

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
We note the linear regression model:
\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}). As shown in section \ref{sectionOLS}, estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\hat{\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model  : the more covariates you have the greater variance you get.
	\item The correlations within the covariates : strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

@@

	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we don't have a system of regressions but one regression on our response variable and a system of sub-regressions (without the response variable). Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss. So we may obtain two kinds of zeros in the marginal model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.
